{ \fontfamily{ptm}\selectfont First Order Nonsmooth Optimization: Algorithm Design, Analysis, Convergence, and Applications }

### 2024-12-12 Notebook LaTEX/Momentum Accelerations/paper3.tex
>w<, changes

                \prod_{i = 0}^{k} \max(1, \rho_{k - 1})

@@

        Recall the stepwise convergence theorem, since the choice of $\gamma, \hat \gamma$ is arbitrary choices for the theorem, we choose $\gamma = \rho_{k - 1} L \alpha_{k - 1}^2, \hat \gamma = L \alpha_k^2$, giving us: 

@@

        The fist inequality is by $\epsilon_k \ge 0$ and the non-negativity of the remaining terms. 

@@

    \begin{proposition}[Abstract similar triangle form]\label{prop:abs-st-form}\;\\
        For all $k\ge 0$, iterates $(w_t, y_t, x_{t + 1}, z_{t + 1})$ satisfies recursively that: 
        \begin{align*}
            y_k &= (1 + \tau_k)^{-1}(v_k + \tau_k x_k),
            \\
            v_{k + 1} &= (1 + \xi_k)^{-1}(v_k + \xi_k y_k) - (1 + \xi_k)^{-1}\delta_k g_k,
            \\
            x_{k + 1} &= y_k - L^{-1} g_k. 
        \end{align*}
        If $1 + \xi_t + \tau_t = L\delta_t$. Then 
        $$
            v_{k + 1} - x_{k + 1} = (1 + \xi_t)^{-1}\tau_t(x_{k + 1} - x_k). 
        $$
        Which makes the algorithm a similar triangle form. 
    \end{proposition}
    \begin{proof}
        We are interested in identifying the conditions required for the sequence of $\xi_t, \tau_t, \delta_t$ such that there exists $\theta_t$ satisfying: 
        \begin{align*}
            x_{t + 1} - z_{t + 1} 
            &= \theta_t(z_{t + 1} - z_t)
        \end{align*}
        To verify, do 
        \begin{align*}
            x_{t + 1} &= 
            (1 + \xi_t)^{-1}(x_t + \xi_t y_t - \delta_t \mathcal G_L(y_t))
            \\
            & \textcolor{gray}{
                x_t = (1 + \tau_t)y_t - \tau_t z_t
            }
            \\
            &= 
            (1 + \xi_t)^{-1}((1 + \tau_t)y_t - \tau_t z_t + \xi_t y_t - \delta_t \mathcal G_L(y_t))
            \\
            &= 
            (1 + \xi_t)^{-1}((1 + \tau_t + \xi_t)y_t - \tau_t z_t - \delta_t \mathcal G_L(y_t))
            \\
            x_{t + 1} - z_{t + 1}
            &= 
            (1 + \xi_t)^{-1}((1 + \tau_t + \xi_t)y_t - \tau_t z_t - \delta_t \mathcal G_L(y_t))
            - (y_t - L^{-1}\mathcal G_Ly_t)
            \\
            &= 
            (1 + \xi_t)^{-1}(\tau_ty_t - \tau_t z_t - \delta_t \mathcal G_L(y_t))
            + L^{-1}\mathcal G_Ly_t
            \\
            &= 
            (1 + \xi_t)^{-1}
            \left(
                \tau_ty_t - \tau_t z_t + (L^{-1}(1 + \xi_t) - \delta_t) \mathcal G_L(y_t)
            \right)
            \\
            &= 
            (1 + \xi_t)^{-1}\tau_t
            \left(
                y_t - z_t + 
                \tau_t^{-1}(L^{-1}(1 + \xi_t) - \delta_t) \mathcal G_L(y_t)
            \right)
        \end{align*}
        The RHS is can be verified through 
        \begin{align*}
            z_{t + 1} - z_t &= 
            y_t - L^{-1}\mathcal G_L(y_t) - z_t
            \\
            &= (y_t - z_t) - L^{-1}\mathcal G_L(y_t). 
        \end{align*}
        It necessitates the condition: 
        \begin{align*}
            \tau_t^{-1}(L^{-1}(1 + \xi_t) - \delta_t) 
            &= - L^{-1}
            \\
            (1 + \xi_t) - L\delta_t
            &= 
            - \tau_t
            \\
            1 + \xi_t + \tau_t
            &=
            L\delta_t. 
        \end{align*}
        Which allows for: 
        \begin{align*}
            x_{t + 1} - z_{t + 1} &= 
            (1 + \xi_t)^{-1}\tau_t
            \left(y_t - z_t - L^{-1}\mathcal G_L(y_t)\right) 
            = 
            (1 + \xi_t)^{-1}\tau_t(z_{t + 1} - z_t). 
        \end{align*}
    \end{proof}

@@

        The coefficient for $(x_{k + 1} - x_k)$ needs some works to formulate it without the parameter $\mu, L$. 
        To do that we have: 
        \begin{align*}
            \frac{(L\alpha_{k + 1} - \mu)(\alpha_k^{-1} - 1)}{L - \mu}
            &= \frac{(L\alpha_{k + 1} - \mu)\alpha_k(1 - \alpha_k)}{\alpha_k^2(L - \mu)}
            \\
            &= 
            \alpha_k(1 - \alpha_k)
            \left(
                \frac{\alpha_k^2(L - \mu)}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= \alpha_k(1 - \alpha_k)
            \left(
                \frac{L\alpha_k^2 - \mu\alpha_k^2}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= 
            \alpha_k(1 - \alpha_k)
            \rho_k\left(
                \frac{L\rho_k\alpha_k^2 - \mu\rho_k\alpha_k^2}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= 
            \rho_k\alpha_k(1 - \alpha_k)
            \left(
                \frac{(L\alpha_{k + 1} - \mu)(\rho_k\alpha_k^2 + \alpha_{k + 1})}
                {L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}.
        \end{align*}
        Going from the left to right on the fourth equality, we used $L\alpha_{k + 1}^2 = (1 - \alpha_{k + 1})\rho_kL\alpha_k^2 + \mu \alpha_{k + 1}$ with 
        \begin{align*}
            L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2 
            &= 
            (1 - \alpha_{k + 1})L \rho_k \alpha_k^2 + \alpha_{k + 1} L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= 
            ((1 - \alpha_{k + 1})L \rho_k \alpha_k^2 + \mu \alpha_{k + 1}) - \mu\alpha_{k + 1} + \alpha_{k + 1} L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= L \alpha_{k + 1}^2 - \mu\alpha_{k + 1} + \alpha_{k + 1}L\rho_k\alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= 
            L\alpha_{k + 1}(\alpha_{k + 1} + \rho_k \alpha_k^2) - \alpha_{k + 1}\mu - \mu \rho_k \alpha_k^2
            \\
            &= (L \alpha_{k + 1} - \mu)(\alpha_{k + 1} + \rho_k \alpha_k^2). 
        \end{align*}
        When $\mu = 0$, things simplify. 
        Consider that $\alpha_{k +1}^2 = (1 - \alpha_{k + 1})\alpha_k^2$. 
        \begin{align*}
            \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}
            &= 
            \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}
            {\rho_k\alpha_{k + 1}\alpha_k^2 + \alpha_{k + 1}^2}
            \\
            &= 
            \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}
            {\rho_k\alpha_{k + 1}\alpha_k^2 + \rho_k(1 - \alpha_{k + 1})\alpha_k^2}
            \\
            &= \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2}
            \\
            &= \alpha_{k + 1}(\alpha_k^{-1} - 1). 
        \end{align*}
    \end{proof}

>>>>

                \prod_{i = 0}^{k - 1} \max(1, \rho_{k})

@@

        Recall the stepwise convergence theorem, since the choice of $\gamma, \hat \gamma$ is arbitrary choices for the theorem, we choose $\gamma = \rho_{k - 1} L \alpha_{k - 1}^2, \hat \gamma = L \alpha_k^2$ at each step, giving us: 

@@

        The first inequality is by $\epsilon_k \ge 0$ and the non-negativity of the remaining terms. 

@@

    \begin{proposition}[Abstract similar triangle form]\label{prop:abs-st-form}\;\\
        For all $k\ge 0$, iterates $(w_t, y_t, x_{t + 1}, z_{t + 1})$ satisfies recursively that: 
        \begin{align*}
            y_k &= (1 + \tau_k)^{-1}(v_k + \tau_k x_k),
            \\
            v_{k + 1} &= (1 + \xi_k)^{-1}(v_k + \xi_k y_k) - (1 + \xi_k)^{-1}\delta_k g_k,
            \\
            x_{k + 1} &= y_k - L^{-1} g_k. 
        \end{align*}
        If $1 + \xi_t + \tau_t = L\delta_t$. Then 
        $$
            v_{k + 1} - x_{k + 1} = (1 + \xi_t)^{-1}\tau_t(x_{k + 1} - x_k). 
        $$
        Which makes the algorithm a similar triangle form. 
    \end{proposition}
    \begin{proof}
        We are interested in identifying the conditions required for the sequence of $\xi_t, \tau_t, \delta_t$ such that there exists $\theta_t$ satisfying: 
        \begin{align*}
            v_{t + 1} - x_{t + 1} 
            &= \theta_t(x_{t + 1} - x_t)
        \end{align*}
        To verify, do 
        \begin{align*}
            v_{t + 1} &= 
            (1 + \xi_t)^{-1}(v_t + \xi_t y_t - \delta_t \mathcal G_L(y_t))
            \\
            & \textcolor{gray}{
                v_t = (1 + \tau_t)y_t - \tau_t x_t
            }
            \\
            &= 
            (1 + \xi_t)^{-1}((1 + \tau_t)y_t - \tau_t x_t + \xi_t y_t - \delta_t \mathcal G_L(y_t))
            \\
            &= 
            (1 + \xi_t)^{-1}((1 + \tau_t + \xi_t)y_t - \tau_t x_t - \delta_t \mathcal G_L(y_t))
            \\
            v_{t + 1} - x_{t + 1}
            &= 
            (1 + \xi_t)^{-1}((1 + \tau_t + \xi_t)y_t - \tau_t x_t - \delta_t \mathcal G_L(y_t))
            - (y_t - L^{-1}\mathcal G_Ly_t)
            \\
            &= 
            (1 + \xi_t)^{-1}(\tau_ty_t - \tau_t x_t - \delta_t \mathcal G_L(y_t))
            + L^{-1}\mathcal G_Ly_t
            \\
            &= 
            (1 + \xi_t)^{-1}
            \left(
                \tau_ty_t - \tau_t x_t + (L^{-1}(1 + \xi_t) - \delta_t) \mathcal G_L(y_t)
            \right)
            \\
            &= 
            (1 + \xi_t)^{-1}\tau_t
            \left(
                y_t - x_t + 
                \tau_t^{-1}(L^{-1}(1 + \xi_t) - \delta_t) \mathcal G_L(y_t)
            \right)
        \end{align*}
        The RHS is can be verified through 
        \begin{align*}
            x_{t + 1} - x_t &= 
            y_t - L^{-1}\mathcal G_L(y_t) - x_t
            \\
            &= (y_t - x_t) - L^{-1}\mathcal G_L(y_t). 
        \end{align*}
        It necessitates the condition: 
        \begin{align*}
            \tau_t^{-1}(L^{-1}(1 + \xi_t) - \delta_t) 
            &= - L^{-1}
            \\
            (1 + \xi_t) - L\delta_t
            &= 
            - \tau_t
            \\
            1 + \xi_t + \tau_t
            &=
            L\delta_t. 
        \end{align*}
        Which allows for: 
        \begin{align*}
            v_{t + 1} - x_{t + 1} &= 
            (1 + \xi_t)^{-1}\tau_t
            \left(y_t - x_t - L^{-1}\mathcal G_L(y_t)\right) 
            = 
            (1 + \xi_t)^{-1}\tau_t(x_{t + 1} - x_t). 
        \end{align*}
    \end{proof}

@@

        The coefficient for $(x_{k + 1} - x_k)$ needs some works to formulate it without the parameter $\mu, L$. 
        To do that we have: 
        \begin{align*}
            \frac{(L\alpha_{k + 1} - \mu)(\alpha_k^{-1} - 1)}{L - \mu}
            &= \frac{(L\alpha_{k + 1} - \mu)\alpha_k(1 - \alpha_k)}{\alpha_k^2(L - \mu)}
            \\
            &= 
            \alpha_k(1 - \alpha_k)
            \left(
                \frac{\alpha_k^2(L - \mu)}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= \alpha_k(1 - \alpha_k)
            \left(
                \frac{L\alpha_k^2 - \mu\alpha_k^2}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= 
            \alpha_k(1 - \alpha_k)
            \rho_k\left(
                \frac{L\rho_k\alpha_k^2 - \mu\rho_k\alpha_k^2}{L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= 
            \rho_k\alpha_k(1 - \alpha_k)
            \left(
                \frac{(L\alpha_{k + 1} - \mu)(\rho_k\alpha_k^2 + \alpha_{k + 1})}
                {L\alpha_{k + 1} - \mu}
            \right)^{-1}
            \\
            &= \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}.
        \end{align*}
        Going from the left to right on the fourth equality, we used $L\alpha_{k + 1}^2 = (1 - \alpha_{k + 1})\rho_kL\alpha_k^2 + \mu \alpha_{k + 1}$ with 
        \begin{align*}
            L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2 
            &= 
            (1 - \alpha_{k + 1})L \rho_k \alpha_k^2 + \alpha_{k + 1} L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= 
            ((1 - \alpha_{k + 1})L \rho_k \alpha_k^2 + \mu \alpha_{k + 1}) - \mu\alpha_{k + 1} + \alpha_{k + 1} L \rho_k \alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= L \alpha_{k + 1}^2 - \mu\alpha_{k + 1} + \alpha_{k + 1}L\rho_k\alpha_k^2 - \mu \rho_k \alpha_k^2
            \\
            &= 
            L\alpha_{k + 1}(\alpha_{k + 1} + \rho_k \alpha_k^2) - \alpha_{k + 1}\mu - \mu \rho_k \alpha_k^2
            \\
            &= (L \alpha_{k + 1} - \mu)(\alpha_{k + 1} + \rho_k \alpha_k^2). 
        \end{align*}
        When $\mu = 0$, things simplify. 
        Consider that $\alpha_{k +1}^2 = (1 - \alpha_{k + 1})\rho_k\alpha_k^2$. 
        \begin{align*}
            \frac{\rho_k\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2 + \alpha_{k + 1}}
            &= 
            \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}
            {\rho_k\alpha_{k + 1}\alpha_k^2 + \alpha_{k + 1}^2}
            \\
            &= 
            \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}
            {\rho_k\alpha_{k + 1}\alpha_k^2 + \rho_k(1 - \alpha_{k + 1})\alpha_k^2}
            \\
            &= \frac{\rho_k\alpha_{k + 1}\alpha_k(1 - \alpha_k)}{\rho_k\alpha_k^2}
            \\
            &= \alpha_{k + 1}(\alpha_k^{-1} - 1). 
        \end{align*}
    \end{proof}

### 2024-12-12 Notebook LaTEX/Proposal/main.tex
>w<, changes

\author{
    \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    The University of British Columbia, Kelowna, BC Canada V1V 1V7. 
    E-mail:  \texttt{alto@mail.ubc.ca}.}~Author Name 1
    \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    The University of British Columbia, Kelowna, BC Canada V1V 1V7. 
    E-mail:  \texttt{shawn.wang@ubc.ca}.}~ and~Author Name 2
}

@@

\begin{abstract} 
    \noindent
    Abstract abstract abstract. 
\end{abstract}
% \noindent{\bfseries 2010 Mathematics Subject Classification:}
% Primary 65K10, 90c25, 90C30; Secondary 65Y20. 
% \noindent{\bfseries Keywords: } Non-convex Optimizations, Proximal Point.

@@

    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had two lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithm. 
        \item Is it possible to attain faster convergence rate without knowledge about the strong convexity of function $f$. 
    \end{enumerate}
    The good news is we have definitive answers for both questions through our own efforts of research. 
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:spectral-momentum} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:spectral-momentum} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par

@@

\section{Nestrov's acceleration in the non-convex case}\label{sec:nes-acc-ncnvx}

>>>>

\author{
    \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    The University of British Columbia, Kelowna, BC Canada V1V 1V7. 
    E-mail:  \texttt{alto@mail.ubc.ca}.}~Hongda Li
    % \thanks{Department of Mathematics, I.K. Barber Faculty of Science,
    % The University of British Columbia, Kelowna, BC Canada V1V 1V7. 
    % E-mail:  \texttt{shawn.wang@ubc.ca}.}~ and~Author Name 2
}

@@

\begin{abstract} 
    \noindent
    The research proposal focuses on the theories and practice in solving nonsmooth optimization. 
    The theme of proposal highlight topics of interests that emphasize the computations and applications aspect of algorithms that exhibits both practical and theoretical importance. 
    We summarize our ongoing research in unifying Nesterov-type accelerated proximal gradient method and proposes our method of spectral momentum. 
    We survey literatures under the topic of Catalyst Meta Acceleration framework used in accelerating variance reduced methods in the settings of Data Science and Machine Learning. 
    We present literatures and progress in topics such as Performance Estimation Problem, Inexact Proximal Point, acceleration without convexity. 
    At the end there is a section summarizing a method we developed for tree species classifications using Sentinel-2 satellite remote sensing data using big data anlaytics by extract spectral signatures of ground vegegation covers.

\end{abstract}
% \noindent{\bfseries 2010 Mathematics Subject Classification:}
% Primary 65K10, 90c25, 90C30; Secondary 65Y20. 
% \noindent{\bfseries Keywords: } Non-convex Optimizations, Proximal Point.

@@

    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had two lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithm. 
        \item Is it possible to attain faster convergence rate without knowledge about the strong convexity of function $f$. 
    \end{enumerate}
    The good news is we have definitive answers for both questions through our own efforts of research. 
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:spectral-momentum} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:spectral-momentum} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par
    Section \ref{sec:catalyst} are results of literatures review in MATH 590. 
    It's based on a series of papers (CITATION HERE) in the topic of Catalyst Meta Acceleration method for First Order Variance Reduced Methods. 
    We will point out potential future direction of research of Catalyst acceleration. 
    \par

@@

\section{Nestrov's acceleration in the non-convex case and momentum in general}\label{sec:nes-acc-ncnvx}
